ANNO: The Anthropological Notation Ontology

### 2023-02-17 paper.tex
More introduction.

% KH: Ich hab hier mal den Text vom Poster als Basis für die Introduction genommen
\section{Introduction}\label{sec:introduction}
Historic and forensic anthropology gives insights into the behaviour, culture and way of life of humans of different epochs in the past.
Osteometry, describes quantitative measurement on the human skeletton.
Anthropology relies on osteometric measurements which has several problems:
Bones can only be investigated in place, which requires transport over far distances.
This causes wear in the bones, and can lead to different measurements.
This can be alleviated with a digital twin...

\section{Ontology Modelling}
TODO: Modellierung der Anthropologie-Domäne (u.a. Motivation, related work, was ist neu).
TODO: Dann könnte man Design-Entscheidungen beschreiben (wie und warum wurden bestimmte Sachverhalte so in der Ontologie modelliert).
\section{Publication}
TODO: Bereitstellung der Ontologie für die Community (OLS etc.). 
\section{Integration into \aw}
TODO: Integration der Ontologie in die Anwendung.
%\subsection{}\label{s1.1}

>>>>

% KH: Ich hab hier mal den Text vom Poster als Basis für die Introduction genommen
\section{Introduction}\label{sec:introduction}
Historic and forensic anthropology gives insights into the behaviour, culture and way of life of humans of different epochs in the past.
Osteometry, describes quantitative measurement on the human skeletton.
Anthropology relies on osteometric measurements which has several problems:
Bones can only be investigated in place, which requires transport over far distances.
This causes wear in the bones, and can lead to different measurements.
This can be alleviated with a digital twin, which can be measured at any location.

\section{Related Work}
The Foundational Model of Anatomy (FMA) ontology \cite{fma} describes concepts of anatomy.
\url{https://bioportal.bioontology.org/ontologies/FMA}

\section{Ontology Modelling}
While there are many textual sources of anthropological systematization, they do not agree in all aspects and there is not a single, formally described, standard.
ANNO links to concepts of the FMA when they exist, but structures them in a different way.
The ontology consists of the following core modules:

\subsection{Bones}

\subsection{Anatomical Landmarks}
Anatomical landmarks are defined either on a bone or an anatomical compound structure.
\subsection{Osteometric Landmarks}

\subsection{Teeth}
\subsection{Measurements}
\subsection{Functions of Sex Determination}
The sex of a specific individual within a population may be estimated using a function on skelettal measurements that is specific to this population.
RDF is not optimized for mathematical formulas so we model those as literals.

\subsection{Height Functions}
\subsection{Landmarks}

TODO: Modellierung der Anthropologie-Domäne (u.a. Motivation, related work, was ist neu).
TODO: Dann könnte man Design-Entscheidungen beschreiben (wie und warum wurden bestimmte Sachverhalte so in der Ontologie modelliert).
\section{Publication}
ANNO is published on \url{https://ols.imise.uni-leipzig.de} as well as \url{https://annosaxfdm.de/ontology/}.
TODO: Bereitstellung der Ontologie für die Community (OLS etc.). 
\section{Integration into \aw}
TODO: Integration der Ontologie in die Anwendung.
%\subsection{}\label{s1.1}
